---
title: "Lab0"
author: "Durley Torres-Marin"
date: "September 2, 2017"
output: html_document
---
```{r}
source("http://www.openintro.org/stat/data/arbuthnot.R")
arbuthnot
dim(arbuthnot)
names(arbuthnot)
```
Exercise 1 : What command would you use to extract just the counts of girls baptized
```{r}
arbuthnot$girls
plot(x = arbuthnot$year, y = arbuthnot$girls)
plot(x = arbuthnot$year, y = arbuthnot$girls, type = "l")
```

Exercise 2: Is there an apparent trend in the number of girls baptized over the years?
How would you describe it? 

The previous graph clearly shows that the girsl baptized after the 1660s are growing, for the period of time between 1640 and 1660 there is a significant drops of girls baptized.

Exercise 3 :Now, make a plot of the proportion of boys over time. What do you see? Tip: If you use the up and down arrow keys, you can scroll through your previous commands, your so-called command history. You can also access it by clicking on the history tab in the upper right panel. This will save you a lot of typing in the future.

```{r}
plot(arbuthnot$year, arbuthnot$boys / (arbuthnot$boys + arbuthnot$girls), type = "l")
```

Over time, I can see a variations year to year of the proportion of boys; however, the proportions over 0.535 means that boys' baptisms went beyond the girls durin 1660s, and there is a significant drops of boys baptisms in the 1700's.
We can see below how boys outnumber girls in each year
```{r}
arbuthnot$boys > arbuthnot$girls
```
Values of either TRUE if that year had more boys than girls, or FALSE if that year did not.

```{r}
source("http://www.openintro.org/stat/data/present.R")
```
1.  What years are included in this data set?  What are the dimensions of the dataframe and what are the variable or column names?
```{r}
present$year
dim(present)
names(present)
```
2.  How do these counts compare to Arbuthnot's?  Are they on a similar scale?
Both data frames Arbuthot and Present are identical in terms of attributes and scale but in the present data frame is clear similar that contain higher numbers of births than Arbuthnot's data frame, respectively
```{r}
present$boys
summary(present$boys)
```
```{r}
present$girls
summary(present$girls)
```
3.  Does Arbuthnot's observation about boys being born in greater proportion than girls hold up in the U.S.?
```{r}
present$boys > present$girls
```
Values of either TRUE if that year had more boys than girls, or FALSE if that year did not.

Make a plot that displays the boy-to-girl ratio for every year in the dataset.  What do you see?
There is a significant drops in the ratio at birth over time. Also radio at birth in boys is greater than girls in the 1940's it went over 1.054 ratio in which more boys were born.

```{r}
plot(x = present$year, y = present$boys/present$girls, type = "l")
```


4. In what year did we see the most total number of births in the U.S.?  


```{r}
present$tot <- present$boys + present$girls
present$tot
```
```{r}
which.max(present$tot)
present$year[22]
```
